---
title: "Visualize Data"
---

## Your Turn 0

Add a setup chunk that loads the tidyverse packages.

# add a chunk: Ctrl+Alt+I  #
```{r setup}
library(tidyverse)
theme_set(theme_bw())
```


```{r}
mpg
```


## Your Turn 1

Run the code on the slide to make a graph. Pay strict attention to spelling, capitalization, and parentheses!

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy))

```

## Your Turn 2

Add `color`, `size`, `alpha`, and `shape` aesthetics to your graph. Experiment.  

```{r}
#This graph plots the relationship between engine size and highway mileage
#and differentiates the class of car by color of the geom point
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, color=class))
#This graph plots the relationship between engine size and highway mileage
#and differentiates the class of car by shape of the geom point
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, shape=class))
#This graph plots the relationship between engine size and highway mileage
#and differentiates the class of car by the transparency of the geom point
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, alpha=class))
#This graph plots the relationship between engine size and highway mileage
#and differentiates the class of car by the size of the geom point
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, size=class))

#More practice
ggplot(data = mpg) +
  theme(legend.position = "none") +
  geom_histogram(aes(x = hwy, binwidth = 1, alpha = 0.3, fill = "#E69F00"))

```

## Help Me

What do `facet_grid()` and `facet_wrap()` do? 

_Facets create subplots that display subsets of the data_

```{r}
# Makes a plot that the commands below will modify
q <- ggplot(mpg) + geom_point(aes(x = displ, y = hwy))

# Plot the relationship between engine size and highway mileage by cylinder in columns
q + facet_grid(. ~ cyl)
# Plot the relationship between engine size and highway mileage by drive type drive type in rows
q + facet_grid(drv ~ .)
# Plot the relationship between engine size and highway mileage by the combinations of drive time and cylinders in a grid
q + facet_grid(drv ~ cyl)
#plot the relationship between engine size and highway mileage by car class type in a rectangular grid.
q + facet_wrap(~ class)


```

## Your Turn 3

Add the black code to your graph. What does it do?

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(displ, hwy, color = class)) +
  labs(title = "Fuel Efficiency by Engine Size", 
       subtitle = "Data facetted by class", 
       x = "Engine Size (displacement in liters)", 
       y = "Fuel Efficiency (MPG)", 
       color = "Class of\nAutomobile", 
       caption = "Data from the EPA")
```

## Your Turn 4

Replace this scatterplot with one that draws boxplots. Use the cheatsheet. Try your best guess.

```{r}
ggplot(mpg, aes(x=class, y=hwy)) + 
  geom_boxplot(mapping = aes(fill=class, notch = TRUE, outlier.colour = "#E69F00", outlier.shape = 8)) + 
  labs(title = "Highway Mileage by Class of Vehicle", 
       x = "Class of Vehilce", 
       y = "Highway Mileage (MPG)")

?geom_boxplot

```

## Your Turn 5

Make a histogram of the `hwy` variable from `mpg`. Hint: do not supply a y variable.

```{r}
ggplot(data = mpg) + 
  geom_histogram(mapping = aes(x=hwy), binwidth=5) + 
  labs(title = "Frequency of Highway Mileage (MPG)", 
       x = "Highway Mileage (MPG)", 
       y = "NUmber of Vehicles")

```

## Your Turn 6

Use the help page for `geom_histogram` to make the bins 2 units wide.

```{r}
ggplot(mpg) +
  geom_histogram(mapping = aes(x=hwy), binwidth=2) + 
  labs(title = "Frequency of Highway Mileage (MPG)", 
       x = "Highway Mileage (MPG)", 
       y = "NUmber of Vehicles")

```

## Your Turn 7

Make a bar chart `class` colored by `class`. Use the help page for `geom_bar` to choose a "color" aesthetic for class.

```{r}
#Fill by class
ggplot(mpg) +
  geom_bar(mapping = aes(x= class, fill=class))

#Fill by drive type
ggplot(mpg) + 
  geom_bar(mapping = aes(x = class, fill=drv))

```

## Quiz

What will this code do?

```{r}
ggplot(mpg) + 
  geom_point(aes(x = displ, y = hwy)) +
  geom_smooth(aes(x = displ, y = hwy))
```
## Global Versus Local

Mappings and data that appear in the `ggplot()` line will apply globally over every layer, while mappings and data what appear in each `geom_` function line will add to or override the global mappings for that layer only. 
```{r}
#Make the same plot as above but instead make the points filled by drive type and the line black and add labels. 

ggplot(data = mpg, mapping = aes(x=displ, y=hwy)) + 
  geom_point(mapping = aes(color=drv)) + 
  geom_smooth(color="black") + 
  labs(title= "Fuel Efficiency (MPG) by Engine Size (displacement in liters)", 
       x= "Engine Size (displacement in L)", 
       y = "Fuel Efficiency (MPG)") +
  scale_color_discrete(name="Drive Type", labels=c("4-Wheel Drive", "Front Wheel Drive", "Rear Wheel Drive"))
```

## Quiz

What is different about this plot? Run the code!

```{r}
p <- ggplot(mpg) + 
  geom_point(aes(displ, hwy)) +
  geom_smooth(aes(displ, hwy))

library(plotly)
ggplotly(p)

ggsave("P", pdf)
```

***

# Take aways

You can use this code template to make thousands of graphs with **ggplot2**.

```{r eval = FALSE}
ggplot(data = <DATA>) +
  <GEOM_FUNCTION>(mapping = aes(<MAPPINGS>))
```